Current status and propsects of R-packages for the design of experiments

### 2021-10-04 paper/rjournal/paper-rjournal.tex
fix commands so it works in html + latex

The critical role of data collection is well captured in the expression
``garbage in, garbage out'' -- in other words, if the collected data is
rubbish then no analysis, however complex it may be, can make something
out of it. Methods for data collection can be dichotomised by the type
of data collected -- namely, experimental or observational -- or
alternatively, categorised into experimental design or survey design.
Survey designs often, although not always, aim to collect observational
data whilst experimental designs exclusively center on experimental
data. This dichotomisation, to a great extent, is seen in CRAN task
views where R-packages in experimental design are in
\ctv{ExperimentalDesign} and R-packages in survey designs are in
\ctv{OfficialStatistics}.

In the CRAN task view of \ctv{ExperimentalDesign}, there are 113 R
packages for experimental design and analysis of data from experiments,
henceforth referred to as ``DoE packages'' in this paper. The sheer
quantity and variation in the output experimental design in the
R-packages are arguably unmatched with any other programming languages,
e.g.~in Python \citep{python}, only a handful of libraries that generate
design of experiment exist (namely \texttt{pyDOE}, \texttt{pyDOE2},
\texttt{dexpy}, \texttt{experimenter} and \texttt{GPdoemd}) with limited
outputs.

\hypertarget{explorative-analysis}{%
\section{Explorative Analysis}\label{explorative-analysis}}

@@

{\centering \includegraphics{paper_files/figure-latex/dlplots-1}

@@

\caption[The above graph shows the total download counts from 2019-03-06 to 2021-03-04 of DoE packages]{The above graph shows the total download counts from 2019-03-06 to 2021-03-04 of DoE packages.}\label{fig:dlplots}
\end{figure}
\end{Schunk}

@@

\begin{center}\includegraphics{paper_files/figure-latex/wordcloud-1} \end{center}

@@

\hypertarget{software-design}{%
\section{Software Design}\label{software-design}}

\hypertarget{propsects-and-discussion}{%
\section{Propsects and Discussion}\label{propsects-and-discussion}}

@@

\begin{table}[h] \centering  
\begin{tabular}[t]{lr}
\toprule
word & n\\
\midrule
optimal design & 10\\
experimental design & 8\\
clinical trial & 5\\
dose finding & 5\\
sequential design & 5\\
block design & 4\\
microarray experiment & 4\\
response surface & 4\\
\bottomrule
\end{tabular} \hspace{1cm} \centering  
\begin{tabular}[t]{lr}
\toprule
word & n\\
\midrule
experimental design & 11\\
optimal design & 10\\
package provide & 7\\
graphical user & 6\\
response surface & 6\\
user interface & 6\\
block design & 5\\
contour plot & 5\\
design based & 5\\
effect model & 5\\
factorial design & 5\\
microarray experiment & 5\\
mixed effect & 5\\
provide function & 5\\
sample size & 5\\
sequential design & 5\\
\bottomrule
\end{tabular} \caption{My tables} \end{table}

\CRANpkg{agricolae}

\hypertarget{helping-info-to-get-started}{%
\subsection{Helping info to get
started}\label{helping-info-to-get-started}}

Introductory section which may include references in parentheses
\citep{R}, or cite a reference such as \citet{R} in the text.

\hypertarget{section-title-in-sentence-case}{%
\subsection{Section title in sentence
case}\label{section-title-in-sentence-case}}

Let's check fi this works Figure @ref(fig:Rlogo).

This section may contain a figure such as Figure \ref{fig:Rlogo}.

\hypertarget{summary}{%
\subsection{Summary}\label{summary}}

This file is only a basic article template. For full details of
\emph{The R Journal} style and information on how to prepare your
article for submission, see the
\href{https://journal.r-project.org/share/author-guide.pdf}{Instructions
for Authors}.

\hypertarget{about-this-format-and-the-r-journal-requirements}{%
\subsubsection{About this format and the R Journal
requirements}\label{about-this-format-and-the-r-journal-requirements}}

\texttt{rticles::rjournal\_article} will help you build the correct
files requirements:

\begin{itemize}
\tightlist
\item
  A R file will be generated automatically using \texttt{knitr::purl} -
  see \url{https://bookdown.org/yihui/rmarkdown-cookbook/purl.html} for
  more information.
\item
  A tex file will be generated from this Rmd file and correctly included
  in \texttt{RJwapper.tex} as expected to build \texttt{RJwrapper.pdf}.
\item
  All figure files will be kept in the default rmarkdown
  \texttt{*\_files} folder. This happens because
  \texttt{keep\_tex\ =\ TRUE} by default in
  \texttt{rticles::rjournal\_article}
\item
  Only the bib filename is to modifed. An example bib file is included
  in the template (\texttt{RJreferences.bib}) and you will have to name
  your bib file as the tex, R, and pdf files.
\end{itemize}

\bibliography{paper.bib}

>>>>

The critical role of data collection is well captured in the expression
``garbage in, garbage out'' -- in other words, if the collected data is
rubbish then no analysis, however complex it may be, can make something
out of it. Methods for data collection can be dichotomised by the type
of data collected -- namely, experimental or observational -- or
alternatively, categorised as experimental design (including
quasi-experimental design) or survey design. This dichotomisation, to a
great extent, is seen in CRAN task views where R-packages in
experimental design are in \ctv{ExperimentalDesign} and R-packages in
survey designs are in \ctv{OfficialStatistics}. Survey designs often,
although not always, aim to collect observational data whilst
experimental designs exclusively center on experimental data. This paper
is concerned with the latter.

In the CRAN task view of \ctv{ExperimentalDesign}, there are 110 R
packages for experimental design and analysis of data from experiments,
henceforth referred to as ``DoE packages'' in this paper. The sheer
quantity and variation in the output experimental design in the
R-packages are arguably unmatched with any other programming languages,
e.g.~in Python \citep{python}, only a handful of libraries that generate
design of experiment exist (namely \texttt{pyDOE}, \texttt{pyDOE2},
\texttt{dexpy}, \texttt{experimenter} and \texttt{GPdoemd}) with limited
outputs. Thus, the study of DoE packages is also revealing into the
current status of the field of experimental design.

The paper is organisd as follows. Section @ref(eda)

\hypertarget{eda}{%
\section{Explorative analysis}\label{eda}}

@@

{\centering \includegraphics{figures/dlplots-1}

@@

\caption[The above graph shows the total download counts from 2019-10-04 to 2021-10-02 of DoE packages]{The above graph shows the total download counts from 2019-10-04 to 2021-10-02 of DoE packages.}\label{fig:dlplots}
\end{figure}
\end{Schunk}

@@

\begin{center}\includegraphics{figures/wordcloud-1} \end{center}

@@

Figure \ref{fig:dlplots} suggests that the \CRANpkg{AlgDesign} followed
by \CRANpkg{agricolae} are far more downloaded than any other DoE
packages. It is worth noting that \CRANpkg{agricolae} imports
\CRANpkg{AlgDesign} thus any download of \CRANpkg{agricolae} likely
results also in a download of \CRANpkg{AlgDesign}.

\hypertarget{software-design}{%
\section{Software design}\label{software-design}}

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

An experiment involves running a number of steps:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Defining the hypothesis or question
\item
  Formulating methods to test hypothesis or answer question
\item
  Planning the experimental protocol:
\end{enumerate}

\begin{itemize}
\tightlist
\item
  Determining experimental resources,
\item
  specifying the data collection procedure,
\item
  constructing the experimental design layout, and
\item
  proposing a model for analysis.
\end{itemize}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\setcounter{enumi}{3}
\tightlist
\item
  Collect data
\item
  Analyse data, interpret result and conclusion
\end{enumerate}

@@

\begin{table}[h] \centering  
\begin{tabular}[t]{lr}
\toprule
word & n\\
\midrule
optimal design & 9\\
experimental design & 8\\
clinical trial & 5\\
dose finding & 5\\
block design & 4\\
microarray experiment & 4\\
response surface & 4\\
sequential design & 4\\
\bottomrule
\end{tabular} \hspace{1cm} \centering  
\begin{tabular}[t]{lr}
\toprule
word & n\\
\midrule
experimental design & 10\\
optimal design & 10\\
package provide & 7\\
graphical user & 6\\
response surface & 6\\
user interface & 6\\
block design & 5\\
contour plot & 5\\
design based & 5\\
effect model & 5\\
factorial design & 5\\
microarray experiment & 5\\
mixed effect & 5\\
provide function & 5\\
sample size & 5\\
\bottomrule
\end{tabular} \caption{My tables} \end{table}

\bibliography{../paper.bib}
